Spatial variation of coronal Fourier power spectra

### 2016-02-03 tex/paper2/paper2.tex
describing how the initial guesses are created

The (geometric) \mFpa\ are shown in Figure \ref{fig:fit171193} for
each of the physical regions considered.  It is clear that, in
general, the \mFpa\ exhibit power-law like characteristics.  The
results for the quiet Sun, loop footpoints and sunspot regions suggest
a \PL\ \PS.  This can be modeled as
\begin{equation}
\label{eqn:pwrlaw}
\mbox{Model $M_{1}$}: P_{1}(\nu) = A\nu^{-n} + C,
\end{equation}
where $\nu$ is the frequency, $A>0$ is a proportionality constant and
$n>0$ is the \PL\ index.  The quantity $C>0$ models the high-frequency
/ low power end of the spectrum where the detection properties of the
detector apparatus are assumed to dominate the observations.

The moss results for 171\AA\ and 193\AA\ are notably different from
the other results.  In comparison to the general trend observed in
other regions, there appears to be excess Fourier power in the range
1-10 mHz.  \cite{2003ApJ...595L..63D} show that {\BF TRACE}
171\AA\ time-series of bright upper transition region emission above
active region plage (also known as moss) is a source of wavelet
power. Periods of significant wavelet power are found in the range 200
to 600 seconds, and typically persist for 4–7 cycles.  Later work by
\cite{2005ApJ...624L..61D} suggested that spicule flux tubes tilted
with respect to the solar surface provide a mechanism by which
oscillatory power from lower in the atmosphere may be channeled to
upper portions of the atmosphere.  This suggests that a second model
$M_{2}$ for the \Fps\ in regions should be considered.  The second
model $M_{2}$ has two contributions to the overall \Fps; a background
power-law of the form of $P_{1}$, and a contribution that is more
localized, $G(\nu)$.  The model is given by
\begin{equation}
\label{eqn:pwrlawbump}
\mbox{Model $M_{2}$}: P_{2}(\nu) = P_{1}(\nu) + G(\nu)
\end{equation}
where the localized contribution is described by
\begin{equation}
\label{eqn:bump}
G(\nu) = \alpha\exp\left[-\frac{(\ln(\nu)-\beta)^{2}}{2\delta^{2}}\right].
\end{equation}
Both these models are fit to the geometrically averaged Fourier power
spectrum in both wavelengths and in all four regions (Figure
\ref{fig:fit171193}).  {\BF Model fitting and parameter estimation are
  implemented using a Bayesian probability based approach.}  This
approach was chosen for its flexibility and consistent error
estimation properties \citep{2013ApJ...769...89I}.  The details of the
model fitting process are described in Appendix \ref{sec:app:ind}.
The position of $G(\nu)$ is restricted by permitting $\beta$ to vary
in the range $0.1-10$ mHz only.  This restriction, which is imposed in
the fitting process (see Appendix \ref{sec:app:ind}), arises from the
observations of significant wavelet power in the period range 200 -
600 seconds as found by \cite{2003ApJ...595L..63D}.  {\BF After this
  fitting process is complete, the models are compared to decide which
  model best describes the data.  The details of the model comparison
  process are described in the following section.}

@@

\subsection{Fitting the data}\label{sec:app:fit}
Section \ref{???} describes two models of the Fourier power spectrum.
These models are fit to the observed Fourier power spectrum as
described in Section \ref{}.

\Fpa\ models are fit by first generating an estimate of the model
parameters.  This estimate is used as an initial guess to a fitting
routine.

Models were fit by varying model parameters to find the
maximum of the log of likelihood of the data given the model.

Fourier power components are exponentially distributed.  Let us assume
that $I_{j}$ is the observed Fourier power at the frequency indexed by
$j$, and $S_{j}$ is the model Fourier power at that same frequency.
The probability of an observed power $I_{j}$ given a model power
$S_{j}$ is given by
\begin{equation}
p_{j}\left(I_{j}|S_{j}\right) =
\frac{1}{S_{j}}\exp\left[-\frac{I_{j}}{S_{j}}\right]
\label{eqn:freqprob}
\end{equation}
If there are $N$ non-zero Fourier frequencies then the likelihood over
all frequencies is
\begin{equation}
L=\prod_{j=1}^{N}p\left(I_{j}|S_{j}\right)
\label{eqn:freqprob:all}
\end{equation}
The
Nelder-Mead method was used to vary model parameters in search of the
maximum of the log likelihood.

The initial guess for model \ref{eqn:} is generated as follows.  An
estimated power law index is

The initial guess for model \ref{eqn:} is generated as follows.

The full code used (and the history of its development) is availble as
a git repository at https://github.com/wafels/rednoise/tree/perpixel.

>>>>

The (geometric) \mFpa\ are shown in Figure \ref{fig:fit171193} for
each of the physical regions considered.  It is clear that, in
general, the \mFpa\ exhibit power-law like characteristics.  The
results for the quiet Sun, loop footpoints and sunspot regions suggest
a \PL\ \PS.  This can be modeled as
\begin{equation}
\label{eqn:pwrlaw}
\mbox{Model $M_{1}$}: P_{1}(\nu) = A_{p}\nu^{-n} + C,
\end{equation}
where $\nu$ is the frequency, $A>0$ is a proportionality constant and
$n>0$ is the \PL\ index.  The quantity $C>0$ models the high-frequency
/ low power end of the spectrum where the detection properties of the
detector apparatus are assumed to dominate the observations.

The moss results for 171\AA\ and 193\AA\ are notably different from
the other results.  In comparison to the general trend observed in
other regions, there appears to be excess Fourier power in the range
1-10 mHz.  \cite{2003ApJ...595L..63D} show that {\BF TRACE}
171\AA\ time-series of bright upper transition region emission above
active region plage (also known as moss) is a source of wavelet
power. Periods of significant wavelet power are found in the range 200
to 600 seconds, and typically persist for 4–7 cycles.  Later work by
\cite{2005ApJ...624L..61D} suggested that spicule flux tubes tilted
with respect to the solar surface provide a mechanism by which
oscillatory power from lower in the atmosphere may be channeled to
upper portions of the atmosphere.  This suggests that a second model
$M_{2}$ for the \Fps\ in regions should be considered.  The second
model $M_{2}$ has two contributions to the overall \Fps; a background
power-law of the form of $P_{1}$, and a contribution that is more
localized, $G(\nu)$.  The model is given by
\begin{equation}
\label{eqn:pwrlawbump}
\mbox{Model $M_{2}$}: P_{2}(\nu) = P_{1}(\nu) + G(\nu)
\end{equation}
where the localized contribution is described by
\begin{equation}
\label{eqn:bump}
G(\nu) = A_{G}\exp\left[-\frac{(\ln(\nu)-\beta)^{2}}{2\delta^{2}}\right].
\end{equation}
Both these models are fit to the geometrically averaged Fourier power
spectrum in both wavelengths and in all four regions (Figure
\ref{fig:fit171193}).  {\BF Model fitting and parameter estimation are
  implemented using a Bayesian probability based approach.}  This
approach was chosen for its flexibility and consistent error
estimation properties \citep{2013ApJ...769...89I}.  The details of the
model fitting process are described in Appendix \ref{sec:app:ind}.
The position of $G(\nu)$ is restricted by permitting $\beta$ to vary
in the range $0.1-10$ mHz only.  This restriction, which is imposed in
the fitting process (see Appendix \ref{sec:app:ind}), arises from the
observations of significant wavelet power in the period range 200 -
600 seconds as found by \cite{2003ApJ...595L..63D}.  {\BF After this
  fitting process is complete, the models are compared to decide which
  model best describes the data.  The details of the model comparison
  process are described in the following section.}

@@

\subsection{Fitting the data}\label{sec:app:fit}
Section \ref{???} describes two models of the Fourier power spectrum.
These models are fit to the observed Fourier power spectrum as
described in Section \ref{}.  \Fpa\ models are fit by first generating
an estimate of the model parameters.  This estimate is used as an
initial guess to a fitting routine.

The initial guess for model \ref{eqn:pwrlaw} is generated as follows.
A power law index $n$ is estimated using the observed Fourier power
spectra at the fifty lowest non-zero frequencies using the estimation
approach detailed in Section \ref{sec:app:est}.  For the six hours of
data studied, this covers the frequency range $0.046\rightarrow2.36
mHz$.  The fifty lowest non-zero frequencies are used as this is
typically the frequency range where the observed \Fps\ is dominated by
a power-law like behavior.  An estimate for the power law amplitude
$A_{p}$ is found be averaging the observed power at the first five
non-zero frequencies, covering the frequency range
$0.046\rightarrow0.28 mHz$.  An estimate for the background constant
value $C$ is defined by taking the exponential of the average of the
logarithm of the observed power over the last fifty Fourier
frequencies, covering the frequency range $39.35\rightarrow41.6 mHz$.
Averaging over the logarithm of the power reduces the chance that the
estimate of $C$ will be strongly biased by the largest power in the
frequency range.

The initial guess for model \ref{eqn:pwrlawbump} is generated as
follows.  Estimates for $n$, $A_{p}$ and $C$ are generated as
described in the previous paragraph.  An estimated power spectrum
$P_{est}$ is then calculated using these estimates and model
\ref{eqn:pwrlaw}.  If there at least ten positive differences between
the observed Fourier power and estimated power spectrum $P_{est}$ in
the frequency range $0.97 \rightarrow 9.31 mHz$, then these positive
differences are used to estimate the parameters of the Gaussian,
equation \ref{eqn:bump}.  The parameters $A_{G}$ and $\beta$ are
estimated as the maximum value of the difference and the logarithm of
the Fourier frequency at which this maximum value occurs,
respectively.  The width $\sigma$ is estimated by calculating the
standard deviation of the positive differences with respect to the
logarithm of the frequency.  If there are less than ten positive
differences in the required frequency range, it is assumed that there
is a very small Gaussian contribution to the observed \Fps.  In this
case, $\ln A_{G}=-100$, $\beta$ is fixed to be half way between the
logarithm lower and upper estimated Gaussian frequency range $0.97
\rightarrow 9.31 mHz$, and $\sigma = 0.1$.

Model fitting was initialized using the initial estimates described
above.  Models were fit by varying the value of model parameters in
order to find the maximum of the log of likelihood of the data given
the model.  If the time-series data are subject to a Gaussian noise,
then the Fourier power components are exponentially distributed.  Let
us assume that $I_{j}$ is the observed Fourier power at the frequency
indexed by $j$, and $S_{j}$ is the model Fourier power at that same
frequency.  The probability of an observed power $I_{j}$ given a model
power $S_{j}$ is given by
\begin{equation}
p_{j}\left(I_{j}|S_{j}\right) =
\frac{1}{S_{j}}\exp\left[-\frac{I_{j}}{S_{j}}\right]
\label{eqn:freqprob}
\end{equation}
If there are $N$ non-zero Fourier frequencies then the likelihood over
all frequencies is
\begin{equation}
L=\prod_{j=1}^{N}p\left(I_{j}|S_{j}\right)
\label{eqn:freqprob:all}
\end{equation}
The Nelder-Mead method was used to vary model parameters in search of
the maximum of the log likelihood.  The full code used (and the
history of its development) is availble as a git repository at
https://github.com/wafels/rednoise/tree/perpixel.
